---
title: "Project Preliminary Report"
author: Matthew Blake and Donovan Roddy
editor_options:
  markdown:
    wrap: 72
  chunk_output_type: console
output: pdf_document
# number_sections: true
---

```{r, echo=FALSE, message=FALSE, warning = FALSE}
library(readr)
library(tidyverse)
library(mosaic)
library(lme4)
nls_clean <- read_csv("~/Stats_316_F24/Project/DR_MB/nls_clean.csv")
nls_clean <- nls_clean %>% mutate(na.prop = rowMeans(is.na(nls_clean))) 

# pivots the data longer. Row observational units are PUBID and Year. 
nls_long<- nls_clean|>
    relocate(PUBID)|>
    pivot_longer(cols = matches("[0-9]{4}$"), names_to = c(".value","year"), names_pattern = "(.*)([0-9]{4}$)" )|>
  relocate(year) 
  
#Need to add an offset for number of years for the DLI variables!
nls_interp <- nls_long|>
  #filter(year <= 2008)|> #drops years after 2008. Sexual health data not recorded 
  mutate(year = as.double(year))|>#turns year from a character to a double
  mutate(Puberty_age = ifelse(
    !is.na(FEMALE_FMP_AGE), FEMALE_FMP_AGE, ifelse(
      !is.na(MALE_PUB_AGE), MALE_PUB_AGE, NA))) |> #creates puberty age variable
  group_by(PUBID)|>
  arrange(year)|>
  filter(any(!is.na(NUM_PARTNERS_SINCE_DLI)))|> # removes people who do not have any recorded sexual partner data
  #mutate(NUM_PARTNERS_SINCE_DLI = ifelse(is.na(NUM_PARTNERS_SINCE_DLI), 0, NUM_PARTNERS_SINCE_DLI))|> # 
  fill(SEX_SINCE_DLI, .direction = "downup")|>
  fill(Avg_GPA, .direction = "downup")|> #fills in AVG_GPA if missing for each person. AVG_GPA is a level 2 variable, so NA values can be filled.
  fill(Region, .direction = "downup")|>#fills in AVG_GPA if missing for each person. AVG_GPA is a level 1 variable.
  fill(Sex, .direction = "downup")|>
  fill(Marriage_Status, .direction = "downup")|>
  fill(Ethnicity, .direction = "downup")|>
  fill(Race, .direction = "downup")|>
  fill(Rural_Urban, .direction = "downup")|>
  fill(NUM_PARTNERS_SINCE_DLI, .direction = "downup")|>
  arrange(PUBID)|>
  mutate(partners = ifelse(NUM_PARTNERS_SINCE_DLI <= 1,1,0))


```

Healthy sexual relationships and their determinants have been a nuanced
and ongoing question among public health professionals. Variation in the
number of sexual partners between genders, ages, and within different
regions has been a topic of particular interest, as sociologists and
sexual health professionals explore trends in sexual relationships over
time. Exploring the role education plays in sexual health has also been
a frequent point of discussion. This project explores whether there is
an association between one’s high school academic performance and both
their self-reported sexual partners over time and their self-reported
frequency of sexual encounters. The data used in this study comes from
the National Longitudinal Study of Youth from 1997 (NLSY97). The survey
looks at 9884 people, born between 1980-1984. The data was collected
through yearly interviews, from 1997 until 2008. The interviews
collected information about employment, demographics, sexual activity,
dating and marriage, family, health trends, and various other
attributes. This study focuses on information related to age when the
respondent experiences puberty, age of first sex encounter, race,
ethnicity, and sex identify. These measures have been reported to be
associated with sexual partners (Kan et al. 2010). Our measure of
academic performance is one’s average GPA, and our dependent variables
of interest are a respondent’s reported number of sexual encounters
since the last interview and a respondent’s reported number of sexual
partners since the last interview.

Kan, Marni L., Yen-hsin Alice Cheng, Nancy S. Landale, and Susan M.
McHale. 2010. “Longitudinal Predictors of Change in Number of Sexual
Partners across Adolescence and Early Adulthood.” Journal of Adolescent
Health 46 (1): 25–31.
<https://doi.org/10.1016/j.jadohealth.2009.05.002>.

## Variable Descriptions:

-   year

    -   Year that response was recorded (AD)

    -   Level 1 variable and observational unit

    -   Collinear with Age and year of birth, so not used in model

-   PUBID

    -   Unique ID for each individual (no units)

    -   Level 2 variable and observation unit

-   Avg_GPA

    -   Average GPA among 7 different high schools surveyed (some
        attended multiple of the high schools)

    -   No units

    -   Level 2 predictor/independent variable

-   Race

    -   Race of respondent

    -   AAPI, American Indian, Black, Other, or White

    -   Level 2 variable

-   Ethnicity

    -   Ethnicity of respondent

    -   Reduced to "Hispanic" or "Not Hispanic" to reduce collinearity
        with "Race"

    -   Level 2 variable

-   Sex

    -   Sex of respondent

    -   Male or Female

    -   Level 2 variable

-   Puberty_Age

    -   Age at which individual started puberty

    -   Unit for age is years since birth

    -   Level 2 variable

-   na.prop

    -   Proportion of responses left blank

    -   Used to partially account for nonrepresentativity of more filled
        in responses

    -   Level 2 variable

-   Region

    -   Region of the US in which respondent lives

    -   US Census region: Northeast, North Central, South, or West

    -   Level 1 variable

-   Rural_Urban

    -   Whether the respondent lives in a rural or urban area

    -   Binary variable: Rural or Urban

    -   Level 1 variable

-   Marriage_Status

    -   Whether the respondent is cohabitating, or has ever been married

    -   Cohabitating, Never Married, Married, Separated, Divorced, or
        Widowed

    -   Level 1 variable

-   SEX_SINCE_DLI

    -   Number of times respondent has had sex since Date of Last
        Interview

    -   Level 1 response / dependent variable

-   NUM_PARTNERS_EVER

    -   Number of sexual partners respondent has ever had

    -   Level 1 response / dependent variable

-   NUM_PARTNERS_SINCE_DLI

    -   Number of sexual partners since Date of Last Interview

    -   Level 1 response / dependent variable

## Data Cleaning Description

The data was originally collected in a wide format, with the
observational unit being the respondent, and the columns indicated the
question and the year in which it was asked. The original data listed
the columns as numbers related to each question. The columns were
renamed to better represent their measures, and the data was pivoted
longer so that each row represented a respondent and their answers to
the questions each year. This will allow us to do multilevel modeling.
Only about one-third of the respondents had data recorded related to
their sexual partners, so the other respondents were dropped. One of the
challenges of this data set is that the number of sexual partners and
number of sexual encounters did not have yearly responses for each
respondent, listing NAs for years in which no response was recorded. To
account for this, the data was interpolated for each respondent since in
theory these numbers would not fluctuate between interview times as the
question asked for the count since that last interview. We plan to do an
offset for modeling, but this was what we did for the convenience of our
exploratory analysis. In addition, for the level 2 variables, such as
average GPA or race and ethnicity, there were also NA values for certain
years. We also interpolated these values as they do not change over
time. Finally, there were some significant outliers in reports of the
number of sexual partners. A couple of respondents reported around 1000,
while the majority were between 0 and 50. To account for this, we simply
removed respondents with extreme values.

## EDA Graphs

```{r, echo=FALSE, message=FALSE, warning = FALSE}
nls_long %>% ggplot(aes(y = SEX_SINCE_DLI, x = Avg_GPA)) + 
  geom_point(size = 0.2, alpha = 0.2) + 
  geom_smooth() + 
  #geom_smooth(method = "lm", color = "red") +
  theme_bw()+
  labs(title = "Sex Frequency versus High School GPA",
         subtitle = "Amount of Sex since Last Interview vs Average GPA",
       y = "Number of Sexual Encounters since Last Interview",
       x = "Average High School GPA")

```

```{r, echo=FALSE, message=FALSE, warning = FALSE}
 nls_interp|>  
  filter(!is.na(Avg_GPA))|>
  filter(Avg_GPA != 0)|>
  mutate(GPA_YN = ifelse(Avg_GPA >= 2.765, "High", "Low"))|>
  group_by(year,Region, Sex, GPA_YN) |>
  summarize(avg_partner = mean(NUM_PARTNERS_SINCE_DLI)) |>
  ggplot(aes(x=year, y=avg_partner, color= as.factor(GPA_YN))) +
  geom_line()+
  facet_wrap(~Region)+
  theme_bw()+  
  labs(title = "Number of Sexual Partners by GPA Achievement",
      subtitle = "Trends in Sexual Partners for those with High versus Low GPAs",        x = "year",         
    y = "Number of Sexual Partners since last Interview", 
    color = "Median GPA",         
    caption = "Data Source: National Longitudinal Study of Youth in 1999")
```

```{r, echo=FALSE, message=FALSE, warning = FALSE}
nls_interp|>
  filter(NUM_PARTNERS_SINCE_DLI < 50)|>
ggplot(aes(x = NUM_PARTNERS_SINCE_DLI, fill = Sex))+
  geom_histogram(bins = 30, alpha = 0.5)+
  facet_wrap(~year)+
  theme_bw()+
  labs(title = "Distribtuions of Sexual Partners Each Year",      x = "Number of Sexual Partners since last Interview",         
    y = "Distribution", 
    color = "Sex",         
    caption = "Data Source: National Longitudinal Study of Youth in 1997")
  

nls_interp|>
  drop_na(NUM_PARTNERS_SINCE_DLI)|>
  filter(NUM_PARTNERS_SINCE_DLI < 50)|>
  group_by(year, Sex)|>
ggplot(aes(x = year, y = NUM_PARTNERS_SINCE_DLI, color = Sex))  +
  geom_smooth(se = FALSE) +
  theme_bw()+  
  labs(title = "Trends in Sexual Partners For Men and Women",      x = "year",         
    y = "Number of Sexual Partners since last Interview", 
    color = "Sex",         
    caption = "Data Source: National Longitudinal Study of Youth in 1997")
```

This chart shows the SEX_SINCE_DLI versus the Average GPA. We see that
there appears to be a slight negative correlation, from the linear fit
line in red. However, the geom_smooth line reveals that for low SSD
numbers, where we have the most data, there is a notable *positive*
correlation between SSD and Avg_GPA, which levels out around 25.

```{r, echo=FALSE, message=FALSE, warning = FALSE}
nls_long %>% ggplot() + geom_density(aes(x = SEX_SINCE_DLI, color = Sex)) + 
  geom_vline(xintercept = 100) +
  geom_vline(xintercept = 200) +
  geom_vline(xintercept = 300) +
  geom_vline(xintercept = 400) +
  geom_vline(xintercept = 500) +
  geom_vline(xintercept = 1000) + #Vertical lines denote multiples of 100, to show that peaks are near round numbers 
labs(title = "Sex since DLI by sex (M/F)",
     x = "Sex since DLI",
     y = "Density",
     color = "Sex")
```

This is a density chart of Sex since DLI, broken up by sex (male and
female). Vertical lines are included at multiples of 100, to show that
the higher values are disproportionately around these round numbers.

## Figure Summaries.

The first figure indicates the relationship between the average GPA and
the number of sexual encounters. The data suggests that initial
increases in the number of sexual encounters seem to be had by those
with higher GPAs, followed by a relatively steady decline. It generally
suggests that those with high numbers of self-reported sexual encounters
had lower GPAs. Figure two looks at the number of sexual partners. It
suggests that generally, those with lower GPAs tend to have more sexual
partners, but the trends in time seem to vary by region. In addition,
the trend in time for those with high GPAs seems to be more consistent.
Figures 3 look at the distribution of sexual partners each year
separated by sex. They seem to suggest that men seem to have more sexual
partners on average than women. They also seems to suggest that of the
indicating having one sexual partner seems to be indicated significantly
more than all all other numbers. In addition, Figure 4 suggests that men
seem to have more partners on average than women, and this trend appears
consistent over time. The final graph indicates the distribution of
sexual encounters for both men and women. The plot suggests that men and
women seem to round when expressing the number of sexual encounters they
have had, as there are peaks at even numbers ending in zero. They tend
to select numbers like 250, 500, and 1000.

## Demographic Summaries

Looking at the number of sexual encounters over time, we find slight
differences in most of the demographic characteristics. Women have more
moderate amounts of sex, whereas men are more represented at the two
extremes. Black Americans are most highly represented among people with
few or no sexual encounters, while white Americans are overrepresented
among those with many. Hispanic Americans, however, have slightly fewer
sexual encounters on average. The West has the most sexual encounters,
while the South has the fewest, on average; urban people have slightly
more sex than rural people, though not significantly. Those who are not
married and not cohabitating have by far the least sex, while those who
are unmarried and cohabitating have the most. Regarding number of sexual
partners, this also seems to vary significantly depending on marriage
type. Separated and divorced people seem to have higher number of sexual
partners, while those who live with others seem to have less. In
addition, those who went through puberty later seemed to have more
sexual partners.

For more information on the code of our analysis, the Read Me found in
our project Folder describes what each of the different Rmd files are
intended for. It goes over our data cleaning code an exploratory
analysis code.

**Next Steps**:

-   Determine what variables, we want to include in our models.

-   Determine what types of multilevel models we want to use.

**Additional Questions**

-   We want to know if, specifically for number of sexual partners, our
    dependent variable should be made Binary. Where 0 indicates if a
    respondent has 1 or fewer sexual partners and 1 indicates if a
    respondent has more than 1 sexual partner.

-   How do we account for the rounding challenges in the self-reported
    sexual encounters?

# Sources

Kan, Marni L., Yen-hsin Alice Cheng, Nancy S. Landale, and Susan M.
McHale. 2010. “Longitudinal Predictors of Change in Number of Sexual
Partners across Adolescence and Early Adulthood.” Journal of Adolescent
Health 46 (1): 25–31.
<https://doi.org/10.1016/j.jadohealth.2009.05.002>.

Our data was collected from the National Longitudinal Survey of Youth--
1997. This survey tracked individuals born from 1980 to 1984.
<https://www.nlsinfo.org/content/cohorts/nlsy97>
