---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}

library(readr)
library(tidyverse)
library(mosaic)
library(lme4)
library(gt)
library(knitr)
library(kableExtra)
nls_clean <- read_csv("Stats_316_F24/Project/DR_MB/nls_clean.csv")
nls_clean <- nls_clean %>% mutate(na.prop = rowMeans(is.na(nls_clean))) #looks at proportions of NA variables

# pivots the data longer. Row observational units are PUBID and Year. 
nls_long<- nls_clean|>
    relocate(PUBID)|>
    pivot_longer(cols = matches("[0-9]{4}$"), names_to = c(".value","year"), names_pattern = "(.*)([0-9]{4}$)" )|>
  relocate(year) 
  
#More data cleaning. Interpolating some values as not all respondents responded to interview questions every year
nls_interp <- nls_long|>
  #filter(year <= 2008)|>#drops years after 2008. Sexual health data not recorded 
  mutate(year = as.double(year))|>#turns year from a character to a double
  mutate(Puberty_age = ifelse(
    !is.na(FEMALE_FMP_AGE), FEMALE_FMP_AGE, ifelse(
      !is.na(MALE_PUB_AGE), MALE_PUB_AGE, NA))) |> #creates puberty age variable
  group_by(PUBID)|>
  arrange(year)|>
  filter(any(!is.na(NUM_PARTNERS_SINCE_DLI)))|> # removes people who do not have any recorded sexual partner data
  #mutate(NUM_PARTNERS_SINCE_DLI = ifelse(is.na(NUM_PARTNERS_SINCE_DLI), 0, NUM_PARTNERS_SINCE_DLI))|> # 
  fill(SEX_SINCE_DLI, .direction = "downup")|>
  fill(Avg_GPA, .direction = "downup")|> #fills in AVG_GPA if missing for each person. AVG_GPA is a level 2 variable, so NA values can be filled.
  fill(Region, .direction = "downup")|>#fills in Region if missing for each person. AVG_GPA is a level 1 variable.
  fill(Sex, .direction = "downup")|>#fills in sex if missing for each person. Sex is a level 1 variable.
  fill(Marriage_Status, .direction = "downup")|> #fills in Marrige if missing for each person. Marriage is a level 1 variable.
  fill(Ethnicity, .direction = "downup")|> #fills in Ethnicity if missing for each person. Ethnicity is a level 2 variable.
  fill(Race, .direction = "downup")|> #fills in Race if missing for each person. Race is a level 2 variable.
  fill(Rural_Urban, .direction = "downup")|> #fills in AVG_GPA if missing for each person. AVG_GPA is a level 1 variable.
  fill(NUM_PARTNERS_SINCE_DLI, .direction = "downup")|>#fills in AVG_GPA if missing for each person. AVG_GPA is a level 1 variable, and a dependent variable.
  arrange(PUBID)


nls_clean %>% ggplot(aes(x = na.prop)) + geom_histogram()

nls_clean %>% summarise(mean_na = mean(na.prop), sd = sd(na.prop))

```


# Exploratory Analysis of sexual partners with various covariates
```{r}


favstats(~NUM_PARTNERS_SINCE_DLI, data = nls_interp)

nls_interp|>
  filter(NUM_PARTNERS_SINCE_DLI < 50)|>
ggplot(aes(x = NUM_PARTNERS_SINCE_DLI))+
  geom_histogram(bins = 30)

#looks distribution of number of sexual partners since last interview by year. 

nls_interp|>
  filter(NUM_PARTNERS_SINCE_DLI < 50)|>
ggplot(aes(x = NUM_PARTNERS_SINCE_DLI))+
  geom_histogram(bins = 30)+
  facet_wrap(~year) 


#distribution of number of sexual partners since last interview by year, by sex We want to see if there are differences in partners by sex
nls_interp|>
  filter(NUM_PARTNERS_SINCE_DLI < 50)|>
ggplot(aes(x = NUM_PARTNERS_SINCE_DLI, fill = Sex))+
  geom_histogram(bins = 30, alpha = 0.5)+
  facet_wrap(~year) 
#distribution of number of sexual partners since last interview by year, by race

nls_interp|>
  filter(NUM_PARTNERS_SINCE_DLI < 50)|>
ggplot(aes(x = NUM_PARTNERS_SINCE_DLI, fill = Race))+
  geom_histogram(bins = 30, alpha = 0.5)+
  facet_wrap(~year) 
#distribution of number of sexual partners since last interview by year, by race
nls_interp|>
  filter(NUM_PARTNERS_SINCE_DLI < 50)|>
ggplot(aes(x = NUM_PARTNERS_SINCE_DLI, fill = Marriage_Status))+
  geom_histogram(bins = 30, alpha = 0.5)
#By Race By Year
nls_interp|>
  filter(NUM_PARTNERS_SINCE_DLI < 50)|>
ggplot(aes(x = NUM_PARTNERS_SINCE_DLI, fill = Marriage_Status))+
  geom_histogram(bins = 30, alpha = 0.5)+
  facet_wrap(~year) 

# exploring trends in sexual partners over time, looking at different sexes. Looking for differenes based on sex and possible interactions
nls_interp|>  
  group_by(year,Region, Sex, Marriage_Status) |>
  summarize(avg_partner = mean(NUM_PARTNERS_SINCE_DLI)) |>
  ggplot() +
  geom_line(aes(x=year, y=avg_partner, color= Sex))+
  facet_wrap(~Marriage_Status)

#trends in sexual partners over time between men and women in different regions. Looking at additional interactions
nls_interp|>  
  group_by(year,Region, Sex) |>
  summarize(avg_partner = mean(NUM_PARTNERS_SINCE_DLI)) |>
  ggplot() +
  geom_line(aes(x=year, y=avg_partner, color= Sex))+
  facet_wrap(~Region)

##trends in sexual partners over time for different marriage statuses in different regions. Looking at possible interactions
nls_interp|>  
  group_by(year,Region, Marriage_Status) |>
  summarize(avg_partner = mean(NUM_PARTNERS_SINCE_DLI)) |>
  ggplot() +
  geom_line(aes(x=year, y=avg_partner, color= Marriage_Status))+
  facet_wrap(~Region)+
  theme_bw()

#trends in sexual partners overtime for those with high compared to low average GPAs

#Distribution of GPAs
nls_interp|>
  filter(Avg_GPA != 0)|> #Zero GPAS taken out
ggplot(aes(x = Avg_GPA))+
  geom_histogram(bins = 30)
# trends in sexual partners over time based on High versus Low GPAs. Looking at if there are differences based on high school academic performance
 nls_interp|>  
  filter(!is.na(Avg_GPA))|>
  filter(Avg_GPA != 0)|> # zero GPAs removed
  mutate(GPA_YN = ifelse(Avg_GPA >= 2.765, "High", "Low"))|>
  group_by(year,Region, Sex, GPA_YN) |>
  summarize(avg_partner = mean(NUM_PARTNERS_SINCE_DLI)) |>
  ggplot(aes(x=year, y=avg_partner, color= as.factor(GPA_YN))) +
  geom_line()+
  theme_bw()

# Sexual partners over time for high compared to low GPAs by different regions. Looking for more interactions
 nls_interp|>  
  filter(!is.na(Avg_GPA))|>
  filter(Avg_GPA != 0)|>
  mutate(GPA_YN = ifelse(Avg_GPA >= 2.765, "High", "Low"))|>
  group_by(year,Region, Sex, GPA_YN) |>
  summarize(avg_partner = mean(NUM_PARTNERS_SINCE_DLI)) |>
  ggplot(aes(x=year, y=avg_partner, color= as.factor(GPA_YN))) +
  geom_line()+
  facet_wrap(~Region)+
  theme_bw()


#  Looking at how marriage status influences sexual partners over time
 
  nls_interp|>  
  group_by(year,Marriage_Status, Sex) |>
  summarize(avg_partner = mean(NUM_PARTNERS_SINCE_DLI)) |>
  ggplot(aes(x=year, y=avg_partner, color= Marriage_Status)) +
  geom_line()+
  theme_bw()
 
 
 # Sexual partners over time for high compared to low GPAs by different Marriage statuses. Looking for interactions
 nls_interp|>  
  filter(!is.na(Avg_GPA))|>
  filter(Avg_GPA != 0)|>
  mutate(GPA_YN = ifelse(Avg_GPA >= 2.765, "High", "Low"))|>
  group_by(year,Marriage_Status, Sex, GPA_YN) |>
  summarize(avg_partner = mean(NUM_PARTNERS_SINCE_DLI)) |>
  ggplot(aes(x=year, y=avg_partner, color= as.factor(GPA_YN))) +
  geom_line()+
  facet_wrap(~Marriage_Status)+
  theme_bw()
 
# Looking at sexual partners for those of different puberty ages.
#intended to explore if age of puberty has any relationship with sexual partners
 favstats(~Puberty_age, data = nls_interp)
 
 nls_interp|>
ggplot(aes(x = Puberty_age))+
  geom_histogram(bins = 15)
 
 nls_interp|>  
  filter(!is.na(Avg_GPA))|>
  filter(Avg_GPA != 0)|>
  filter(!is.na(Puberty_age))|>
  mutate(Pub_Late = ifelse(Puberty_age >= 12, "Late", "Early"))|>
  group_by(year,Pub_Late, Sex) |>
  summarize(avg_partner = mean(NUM_PARTNERS_SINCE_DLI)) |>
  ggplot(aes(x=year, y=avg_partner, color= as.factor(Pub_Late))) +
  geom_line()+
  theme_bw()
 
```

```{r}
#More analysis of the effect GPA has on sexual partners. First plot removes observations with partners greater than 1000 as these are unusually high.
nls_interp |> 
  filter(NUM_PARTNERS_SINCE_DLI <= 50)|>
  ggplot(aes(x = NUM_PARTNERS_SINCE_DLI, y = Avg_GPA)) + geom_point(size = 0.2, alpha = 0.2) + geom_smooth() + geom_smooth(method = "lm", color = "red")+
  theme_bw()


nls_interp |> 
  filter(NUM_PARTNERS_SINCE_DLI <= 250)|>
  ggplot(aes(x = NUM_PARTNERS_SINCE_DLI, y = Avg_GPA)) + geom_point(size = 0.2, alpha = 0.2) + geom_smooth() + geom_smooth(method = "lm", color = "red")+
  theme_bw()
```


#Models
```{r}
nls_interp2 <- nls_long|>
  #filter(year <= 2008)|>#drops years after 2008. Sexual health data not recorded 
  mutate(year = as.double(year))|>#turns year from a character to a double
  mutate(Puberty_age = ifelse(
    !is.na(FEMALE_FMP_AGE), FEMALE_FMP_AGE, ifelse(
      !is.na(MALE_PUB_AGE), MALE_PUB_AGE, NA))) |> #creates puberty age variable
  drop_na(NUM_PARTNERS_SINCE_DLI)|>
  group_by(PUBID)|>
  mutate(interview_dist = year- lag(year),
         interview_dist = ifelse(is.na(interview_dist),1,0),
         year_center = year - 1997,
         GPA_center = Avg_GPA - 2.542928,
         relationship = ifelse(Marriage_Status == "Divorced", "Not Cohabit", ifelse(Marriage_Status =="Never Married","Not Cohabit" ,ifelse(Marriage_Status =="Separated","Not Cohabit", ifelse(Marriage_Status =="Widowed","Not Cohabit","Cohabit")))),
        partners = ifelse(NUM_PARTNERS_SINCE_DLI <= 1, 0,1)
         )
favstats(~GPA_center, data = nls_interp2)


 nls_interp2|>  
  drop_na(relationship)|>
  group_by(year,relationship, Sex) |>
  summarize(avg_partner = mean(NUM_PARTNERS_SINCE_DLI)) |>
  ggplot(aes(x=year, y=avg_partner, color= relationship)) +
  geom_line()+
  theme_bw()

#
nls_interp|>
  group_by(PUBID)|>
  summarize(event = mean(partners, na.rm = TRUE))

Partners1 <- glmer(NUM_PARTNERS_SINCE_DLI ~  year_center + (1 | PUBID), data = nls_interp2, family = poisson)
summary(Partners1)

Partners2 <- glmer(NUM_PARTNERS_SINCE_DLI ~   + year_center + Avg_GPA + (1 | PUBID), data = nls_interp2, family = poisson)

summary(Partners2)

Partners3 <- glmer(NUM_PARTNERS_SINCE_DLI ~  year_center + Avg_GPA + (1 | PUBID), data = nls_interp2, family = poisson, offset = interview_dist)
summary(Partners3)

Partners4 <- glmer(NUM_PARTNERS_SINCE_DLI ~  year_center + Avg_GPA + relationship + (1 | PUBID), data = nls_interp2, family = poisson, offset = interview_dist )

summary(Partners4)

Partners5 <- glmer(NUM_PARTNERS_SINCE_DLI ~  Sex + year_center + Avg_GPA + relationship + (1 | PUBID), data = nls_interp2, family = poisson, offset = interview_dist)

summary(Partners5)

Partners6 <-  glmer(NUM_PARTNERS_SINCE_DLI ~  Region + Sex + year_center + Avg_GPA + relationship + (1 | PUBID), data = nls_interp2, family = poisson, offset = interview_dist)
summary(Partners6)

Partners7 <-  glmer(NUM_PARTNERS_SINCE_DLI ~ Sex + GPA_center + year_center + GPA_center:year_center + relationship  + (year_center | PUBID), data = nls_interp2, family = poisson, offset = interview_dist)
summary(Partners7)

favstats(~NUM_PARTNERS_SINCE_DLI, data = nls_interp2)


exp(-0.081636)

exp(0.055117)

exp( 0.315933 )
```


```{r Libraries and Dataset Setup 2, echo = FALSE, include = FALSE}
# pivots the data longer. Row observational units are PUBID and Year. 
nls_clean <- read.csv("~/Stats_316_F24/Project/DR_MB/nls_clean.csv")
nls_long <- nls_clean|>
    relocate(PUBID)|>
    pivot_longer(cols = matches("[0-9]{4}$"), names_to = c(".value","year"), names_pattern = "(.*)([0-9]{4}$)" )|>
  relocate(year) |>
   mutate(year = as.numeric(year),
         year_since_1997 = year - 1997, #Centering year
         age = year - Birth_Year, #creating Age
         Puberty_age = ifelse(!is.na(MALE_PUB_AGE), MALE_PUB_AGE, 
                              ifelse(!is.na(FEMALE_FMP_AGE), FEMALE_FMP_AGE, NA)))|> #Combining puberty age variables
#select(1:10, 29:36, 38, everything())drop_na(NUM_PARTNERS_SINCE_DLI)|>
  group_by(PUBID)|>
  mutate(interview_dist = year- lag(year),
         interview_dist = ifelse(is.na(interview_dist),1,0),
         year_center = year - 1997,
         relationship = ifelse(Marriage_Status == "Divorced", "Not Cohabit", ifelse(Marriage_Status =="Never Married","Not Cohabit" ,ifelse(Marriage_Status =="Separated","Not Cohabit", ifelse(Marriage_Status =="Widowed","Not Cohabit","Cohabit")))), #Condenses Marriage Status
        partners = ifelse(NUM_PARTNERS_SINCE_DLI <= 1, 0,1) 
         )

nls_interp <- nls_long|>
  #filter(year <= 2008)|>#drops years after 2008. Sexual health data not recorded 
  mutate(year = as.double(year))|>#turns year from a character to a double
  mutate(Puberty_age = ifelse(
    !is.na(FEMALE_FMP_AGE), FEMALE_FMP_AGE, ifelse(
      !is.na(MALE_PUB_AGE), MALE_PUB_AGE, NA))) |> #creates puberty age variable
  group_by(PUBID)|>
  mutate(age = year - Birth_Year) %>%
  arrange(year)|>
  filter(any(!is.na(NUM_PARTNERS_SINCE_DLI)))|> # removes people who do not have any recorded sexual partner data
  ungroup() %>%
  #mutate(NUM_PARTNERS_SINCE_DLI = ifelse(is.na(NUM_PARTNERS_SINCE_DLI), 0, NUM_PARTNERS_SINCE_DLI))|> # 
  fill(SEX_SINCE_DLI, .direction = "downup")|>
  fill(Avg_GPA, .direction = "downup")|> #fills in AVG_GPA if missing for each person. AVG_GPA is a level 2 variable, so NA values can be filled.
  fill(Region, .direction = "downup")|>#fills in Region if missing for each person. AVG_GPA is a level 1 variable.
  fill(Sex, .direction = "downup")|>#fills in sex if missing for each person. Sex is a level 1 variable.
  fill(Marriage_Status, .direction = "downup")|> #fills in Marrige if missing for each person. Marriage is a level 1 variable.
  fill(Ethnicity, .direction = "downup")|> #fills in Ethnicity if missing for each person. Ethnicity is a level 2 variable.
  fill(Race, .direction = "downup")|> #fills in Race if missing for each person. Race is a level 2 variable.
  fill(Rural_Urban, .direction = "downup")|> #fills in AVG_GPA if missing for each person. AVG_GPA is a level 1 variable.
  fill(NUM_PARTNERS_SINCE_DLI, .direction = "downup")|>#fills in AVG_GPA if missing for each person. AVG_GPA is a level 1 variable, and a dependent variable.
  arrange(PUBID)


nls_SSD_offset <- nls_interp %>% filter(!is.na(SEX_SINCE_DLI) & !is.na(Avg_GPA) & !is.na(age) & !is.na(Marriage_Status))  %>%
  drop_na(SEX_SINCE_DLI)%>%
  group_by(PUBID) %>%
  mutate(interview_dist = year- lag(year),
         interview_dist = ifelse(is.na(interview_dist), 1, interview_dist)) %>%
  ungroup()

nls_interp2 <- nls_long|>
  #filter(year <= 2008)|>#drops years after 2008. Sexual health data not recorded 
  mutate(year = as.double(year))|>#turns year from a character to a double
  mutate(Puberty_age = ifelse(
    !is.na(FEMALE_FMP_AGE), FEMALE_FMP_AGE, ifelse(
      !is.na(MALE_PUB_AGE), MALE_PUB_AGE, NA))) |> #creates puberty age variable
  drop_na(NUM_PARTNERS_SINCE_DLI)|>
  group_by(PUBID)|>
  mutate(interview_dist = year- lag(year),
         interview_dist = ifelse(is.na(interview_dist), 1, interview_dist), #creates interview time variable
         year_center = year - 1997,
         age_center = age - 20, 
         relationship = ifelse(Marriage_Status == "Divorced", "Not Cohabit", ifelse(Marriage_Status =="Never Married","Not Cohabit" ,ifelse(Marriage_Status =="Separated","Not Cohabit", ifelse(Marriage_Status =="Widowed","Not Cohabit","Cohabit")))),
        partners = ifelse(NUM_PARTNERS_SINCE_DLI <= 1, 0, 1),
        GPA_center = Avg_GPA - summarise(nls_interp %>% ungroup(), mean(Avg_GPA, na.rm = TRUE))[[1]]) |>
  ungroup()

nls_interp2|>
  distinct(interview_dist)

```

```{r}
nls_interp2 |> 
  #filter(NUM_PARTNERS_SINCE_DLI <= 250)|>
  ggplot(aes(x = SEX_SINCE_DLI, y = Avg_GPA)) + geom_point(size = 0.2, alpha = 0.2) + geom_smooth() + geom_smooth(method = "lm", color = "red")+
  theme_bw()

favstats(~Avg_GPA, data = nls_interp2)
```


Level 1:

$$\log(\frac{\lambda_{ij}}{interview\_dist_{ij}}) = a_i + b_iYear_{ij} + d_iRelationship_{ij}$$
Level 2:

$$a_i = \alpha_0 + \alpha_1Sex_i + \alpha_2Avg\_GPA_i + u_i$$

$$b_i = \beta_0 + \beta_1Avg\_GPA_i +v_i$$
$$c_i = \gamma_0$$
$$d_i = \delta_0$$
Composite Model
$$\log(\frac{\lambda_{ij}}{interview\_dist_{ij}}) =  \alpha_0 + \alpha_1Sex_i + \alpha_2Avg\_GPA_i + u_i + \beta_0Year_{ij} + \beta_1Avg\_GPA_i*Year_{ij} + v_iYear + \delta_0Relationship_{ij}$$

$\alpha_0$: 

$\alpha_1$ On average, after adjusting for the year, average GPA, Region, and whether they were co-habitating with a partner, Men have 0.2 more sexual partners than women. 

$ \alpha_2$ There is statistically significant evidence that on average, after adjusting for the year, Sex, Region, and whether they were co-habitating with a partner, unit increase in an individuals average GPA is associated with around 1 less reported sexual partner per interview distance. 

$ \beta_0$: On average, after adjusting for the sex, average GPA, Region, and whether they were co-habitating with a partner, the average respondent is expected to report about 1 additional sexual partner each year 

$\gamma_0$: 

$\delta_0$: On average, after adjusting for time, the sex, average GPA, Region the average respondent who is not co-habitating with a partner is expected to report about 1.37  more sexual partners than those who are co-habitating.

$\sigma_u$ The respondent to respondent St. Deviation in number of sexual partners for those in the women in the north central who were not cohabitating with a partner is  1.4438   

$\sigma_v = $ the respondent to respondent St. Deviation of the slope on year is 0.1871, after adjusting for their age, average GPA, sex, region, and if there were co-habitating with a partner. 

$\rho_{uv} = -0.81:$ The correlation between the random variables $u_i$ and $v_i$ is -0.81. This means that in general, a higher variance in number of sexual partners early in one's life will tend to decrease as a person's age increases, and a low variance in number of sexual encounters early in one's life will tend to increase. 

```{r}

summary(Partners7)

fixed_effects <- data.frame(
  variable = c("(Intercept)", "RegionNortheast", "RegionSouth", "RegionWest", 
           "SexMale", "year_center", "Avg_GPA", "relationshipNot Cohabit"),
  estimate = c(-0.047040, -0.352494, 0.011709, 0.194583, 0.204055,  0.055117, 
              -0.081636, 0.315933),
  std_error = c(0.056066, 0.038445, 0.030039, 0.032369, 0.030245, 0.003924, 
                0.014716, 0.013116),
  z_value = c(-0.839, -9.169, 0.390, 6.011, 6.747, 14.046,-5.547, 24.087 )
)

# Random effects and residuals
random_effects <- data.frame(
  group = c("PUBID  (Intercept)", "year_center"),
  variance = c(2.08453, 0.03499),
  std_dev = c(1.4438, 0.1871),
Corr = -0.84)

# Table for fixed effects
fixed_effects_table <- fixed_effects %>%
  gt() %>%
  tab_header(title = "MLM Poisson Fixed Effects") %>%
  cols_label(
    variable = "Effect",
    estimate = "Estimate",
    std_error = "Standard Error",
    z_value = "t-value"
  ) %>%
  fmt_number(
    columns = c(estimate, std_error, z_value), 
    decimals = 3
  )

# Table for random effects
random_effects_table <- random_effects %>%
  gt() %>%
  tab_header(title = "Random Effects for Linear Mixed Model") %>%
  cols_label(
    group = "Group",
    variance = "Variance",
    std_dev = "Standard Deviation"
  ) %>%
  fmt_number(
    columns = c(variance, std_dev), 
    decimals = 3
  )
kable(summary(Partners7))
kable(fixed_effects_table)
kable(random_effects_table)
```


### 2.2.1 Frequency of Sex Since Last Interview

The first model we tried used only Avg_GPA as a predictor for number of sexual encounters. 

Sex Encounters Naive Model 0: 

$$\log(\frac{\lambda_{ij}}{interview\_dist_{ij}}) = \alpha_0 + \alpha_1Avg\_GPA_i + u_i$$

```{r, Sex Count: Naive model, echo = FALSE, cache = TRUE, warning = FALSE}
nls_fil1 <-nls_interp %>% filter(!is.na(SEX_SINCE_DLI) & !is.na(Avg_GPA) & !is.na(age) & !is.na(Marriage_Status))
nls_SSD_offset <- nls_fil1  %>%
  drop_na(SEX_SINCE_DLI)%>%
  group_by(PUBID) %>%
  mutate(interview_dist = year- lag(year),
         interview_dist = ifelse(is.na(interview_dist), 1, interview_dist))

#glmer0 <-glmer(SEX_SINCE_DLI ~ (1 | PUBID), data = nls_SSD_offset, family = poisson, offset = interview_dist)

#saveRDS(glmer0, "~/Stats_316_F24/Project/DR_MB/glmer0.rds")
glmer0 <-read_rds("~/Stats_316_F24/Project/DR_MB/glmer0.rds")


#glmer1 <-glmer(SEX_SINCE_DLI ~  Avg_GPA + (1 | PUBID), data = nls_SSD_offset, family = poisson, offset = interview_dist)

#saveRDS(glmer1, "~/Stats_316_F24/Project/DR_MB/glmer1.rds")
glmer1 <-read_rds("~/Stats_316_F24/Project/DR_MB/glmer1.rds")


#anova(glmer0, glmer1) %>% as.data.frame() %>% kable()

summary(glmer1)$coefficients %>%
  as.data.frame() %>% rownames_to_column(var = "Variables") %>%
  as_tibble() %>%
  kable(escape = TRUE, caption = "Naive Model 0 Summary")
#Avg_GPA not significant by itself
```

This test shows that average HS GPA is not significant by itself. This
gives a preliminary answer to our question "Is Avg GPA a significant
predictor of number of sexual encounters?"

However, this basic model is a significant oversimplification of the
picture, because people change over time. Our second model will account
for the age and marriage / cohabitation status at the time of interview.
The addition of these level one variables will hopefully allow us to
isolate any impact Average GPA might have.


Composite Model 1:

$$\log(\frac{\lambda_{ij}}{interview\_dist_{ij}}) = \alpha_0 + \alpha_1Avg\_GPA_i + \delta_0Relationship_{ij} + u_i + \beta_0Age_{ij} + \beta_1Avg\_GPA*Age_{ij} +v_iAge_{ij}$$

```{r, Sex Count: Final Model, echo = FALSE, cache = TRUE, warning = FALSE}
#Commenting out the model creation code because it was taking a really long time to knit. To check that model produces 

#glmer2 <-glmer(SEX_SINCE_DLI ~ Avg_GPA:age + Avg_GPA + age + Marriage_Status + (age | PUBID), data = nls_interp2, family = poisson, offset = interview_dist)

#glmer3 <-glmer(SEX_SINCE_DLI ~ Sex + GPA_center + age_center + GPA_center:age_center + relationship + (age_center | PUBID), data = nls_interp2, family = poisson, offset = interview_dist)

#
#saveRDS(glmer3, "~/Stats_316_F24/Project/DR_MB/glmer2.rds")
frequency <-read_rds("~/Stats_316_F24/Project/DR_MB/glmer2.rds")
#summary(glmer3)
#summary(glmer2)


summary(frequency)$coefficients %>% as.data.frame()  %>%
  kable(escape = TRUE, caption = "Composite Model 1 Fixed Effects Summary")

frequency %>% VarCorr() %>% as.data.frame() %>%
  kable(escape = TRUE, caption = "Composite Model 1 Random Effects Summary")

#install.packages("stargazer")
#library(stargazer)

#stargazer(glmer2, type = "text", style = "default", summary = FALSE, title = "Model Results- Sex Counts", digits = 3)

```

After adjusting for biological sex and if a respondent was cohabitating
with a partner, each additional year, for individuals with the an
average high school GPA value, we expect the sexual frequency per number
of years since their last interview to decrease by a factor of 0.98.
This effect is **statistically significant**. In addition, there is
**statistically significant** evidence that those who are age 17
experience a 0.62 times decline in sexual frequency for each additional
point in GPA compared to a 0.7 times decline for those who are 23. In
addition, after adjusting for sex, relationship status, and GPA, the
respondent-to-respondent standard deviation in the number of sexual
partners per year since the last interview as they age is 0.3919.
Moreover, on average, after adjusting for the GPA, age, and whether they
were cohabitating, men had lower likelihood in sexual frequency per
number of years since their last interview by a factor of 0.8 than
women. Finally, after adjusting for sex, GPA, and Age, those who were
not cohabitating had a sexual frequency per number of years since their
last interview that was lower by a factor of 0.57 than those who were
cohabitating. This is also statistically significant.

### 2.2.2 Number of Sexual Partners Since Last Interview

Composite Model 2: 
$$\log(\frac{\lambda_{ij}}{interview\_dist_{ij}}) =  \alpha_0 + \alpha_1Sex_i +\delta_0Relationship_{ij} + \alpha_2Avg\_GPA_i + u_i +  \beta_0Age_{ij} + \beta_1Avg\_GPA_i*Age_{ij}+v_iAge{ij}$$

```{r, Partner Count: Model, echo = FALSE, cache = TRUE, warning = FALSE}

#Partners7 <-  glmer(NUM_PARTNERS_SINCE_DLI ~ Sex + GPA_center + age_center + GPA_center:age_center + relationship + (age_center | PUBID), data = nls_interp2, family = poisson, offset = interview_dist)
#summary(Partners7)

#saveRDS(Partners7, "~/Stats_316_F24/Project/DR_MB/partners7.rds")
Partners <-read_rds("~/Stats_316_F24/Project/DR_MB/partners7.rds")

#favstats(~age, data =nls_interp2)
#favstats(~interview_dist, data = nls_interp2)

summary(Partners)$coefficients %>% as.data.frame()  %>%  kable(escape = TRUE, caption = "Composite Model 2 Fixed Effects Summary")

Partners %>% VarCorr() %>% as.data.frame()  %>% kable(escape = TRUE, caption = "Composite Model 2 Random Effects Summary") 
 
  
#stargazer(Partners7, type = "text", style = "default", summary = FALSE, title = "Model Results- Partner Counts", digits = 3) 


```

After adjusting for biological sex and if a respondent was cohabitating
with a partner, each additional year, for individuals with the an
average high school GPA value, we expect the number of sexual partners
per number of years since their last interview to increase by a factor
of 1.001. This effect is **not statistically significant**.
Moreover,those who are age 17 are expected to have a 1.024263 times
increase in sexual frequency for each additional point in GPA compared
to a 1.032966 times increase for those who are 23. This effect is also
**not statistically significant**. In addition, after adjusting for sex,
relationship status, and GPA, the respondent-to-respondent standard
deviation in the number of sexual partners per year since the last
interview is 0.2785. Moreover, on average, after adjusting for the GPA,
age, and whether they were cohabitating, men had 1.005 times more sexual
partners per time since last interviewed than women. This is also not
statistically significant.



```{r,fig.height=5, fig.width=10, echo = FALSE, warning = FALSE, message = FALSE}
#Frequency by GPA and Age
nls_SSD_offset|>
    mutate(GPA_cuts = cut(Avg_GPA, breaks = c(-Inf, 1.98,  3.24, Inf), 
  labels = c("low", "middle", "high")))%>%
  filter(!is.na(NUM_PARTNERS_SINCE_DLI)) %>%
  group_by(age,GPA_cuts) |>
  summarize(avg_sex = mean(NUM_PARTNERS_SINCE_DLI / interview_dist, na.rm = TRUE)) |>
  ggplot() +
  geom_smooth(aes(x= age, y=avg_sex, color= GPA_cuts), se = FALSE) +
  labs(title = "Average Number of Sexual Partners by GPA", 
       x = "Age",
       y = "Number of Sexual Partners Since Last Interview",
       caption = "Data Source: National Longitudinal Study of Youth in 1997") +
  theme(plot.caption = element_text(hjust = 0.5))+
  theme_bw()


nls_SSD_offset|>
    mutate(GPA_cuts = cut(Avg_GPA, breaks = c(-Inf, 1.98,  3.24, Inf), 
  labels = c("low", "middle", "high")))%>%
  filter(!is.na(SEX_SINCE_DLI)) %>%
  group_by(age,GPA_cuts) |>
  summarize(avg_sex = mean(SEX_SINCE_DLI / interview_dist, na.rm = TRUE)) |>
  ggplot() +
  geom_smooth(aes(x= age, y=avg_sex, color= GPA_cuts), se = FALSE) +
  labs(title = "Average Sexual Frequency by GPA", 
       x = "Age",
       y = "Reported Sexual Frequency Since Last Interview",
       caption = "Data Source: National Longitudinal Study of Youth in 1997") +
  theme(plot.caption = element_text(hjust = 0.5))+
  theme_bw()
```

$$\log(\frac{\lambda_{ij}}{interview\_dist_{ij}}) = \beta_0 + \beta_0Age_{ij} + \beta_1Age_{ij}^2 + \alpha_0Sex_i + \alpha_1Sex_iAge_{ij} + \alpha_2Sex_iAge_{ij}^2+\delta_0Relationship_{ij} + \\ \delta_1Relationship_{ij}Age_{ij} + \delta_2Relationship_{ij}Age_{ij}^2+ \gamma_0Avg_GPA_i + \gamma_1Avg_GPA_i*Age_{ij}+ \gamma_1Avg_GPA_i*Age_{ij}^2+ u_i $$
$$\log(\frac{\lambda_{ij}}{interview\_dist_{ij}}) =  \alpha_0 + \beta_0Sex_i + \gamma_0Avg\_GPA_i + \delta_0Relationship_{ij} + \\ (\alpha_1 + \beta_1 Sex_i + \gamma_1Avg\_GPA_i + \delta_1Relationship_{ij}) * Age_{ij} + \\
(\alpha_2 + \beta_2 Sex_i + \gamma_2Avg\_GPA_i + \delta_2Relationship_{ij}) * Age_{ij}^2$$